The Evolution of Phenotypically Invariant Genetic Networks

### 2021-08-11 review-responses2.tex

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

We are writing to submit a revision of our manuscript, ``{System drift and speciation}'' (ID 21-0262), for your review,
and would like to thank you and the reviewers for your suggestions, as they have helped us improve our manuscript's clairty.
We also would like to thank you and the reviewers for the enthusiasm and positive feedback; we are happy that our manuscript was seen as interesting and relevant.

Both reviewers made comments about the section on Haldane's rule, of which we have revised to address their concerns. 
We also made an effort to provide more context for our manuscript by citing the relevant literature, as requested by Reviewer 1. 
To address some of Reviewer 2's concerns, we added more discussion and detail pertaining to system theory and its biological interpretation.

We hope that by making these revisions you will find this manuscript fit for publication.

%We are writing to submit a revision of our manuscript, ``{System drift and speciation}'' (ID 18-0316), for your review. 
%We sincerely apologize for the delayed resubmission but hope that you will find the
%revised manuscript significantly improved.
%
%We thank you and the two reviewers for providing constructive suggestions 
%and feedback that has helped to improve the manuscript. We have followed 
%the suggestions closely and hope to have fully addressed all the concerns and 
%suggestions as detailed in our point-by-point response below. 
%
%In our original submission, we studied the effect of gene regulatory network evolution on speciation,
%first by applying linear system theory to characterize the large space of phenotypically equivalent network organizations,
%and second by using quantitative genetics to show that neutral genetic drift on this network space can lead to speciation over plausible timescales. 
%
%We agree with the reviewers that the combination of these two approaches, while interesting,
%was not seamless and that the quantitative genetics section required further development, 
%and thus have followed the recommendation of both the Editor and Associate Editor,
%by breaking the manuscript into two papers.
%One paper -- the present revision -- covers the application of system theory to regulatory networks,
%and the other (which we plan to resubmit soon) will cover the quantitative genetics.
%In addition to the revised text, we are submitting as supplementary files,
%a detailed response to the reviews and a color-coded diff to make it easy to see what has been removed.
%We are happy with the result -- as predicted by the reviewers, the resulting paper stands alone
%without the quantitative genetics results.
%
%We hope that by focusing this paper on system theory, we have clarified our results, which include a mathematical description of phenotypically equivalent network space 
%and its implications for speciation under neutral genetic drift. These results suggest that the space of phenotypically equivalent network organizations is substantial,
%and often not reproductively compatible, leaving many opportunities for independently evolving populations to become reproductively incompatible.
%The linear system framework applied in this manuscript explicitly describes possible molecular pathways and naturally predicts Haldane's rule.
%
%We would like to note that the previous version of this paper as a preprint has since received 10 citations as well as attracted
%substantial positive attention as a preprint (e.g., see tweets at
%\url{https://www.biorxiv.org/content/10.1101/231209v2}). Furthermore, this paper has been downloaded more than 2,326 times (\url{https://rxivist.org/papers/24911}),
%making it the 155$^{\text{th}}$ most downloaded evolutionary biology paper available on the \emph{bioRxiv}, as of this writing.
%

### 2021-08-11 review-responses2.tex

\reply{
  Yes, that is correct. On average, $F_2$s will have lower fitness than $F_1$s. As the reviewer points out, some $F_2$s could have higher fitnesses than $F_1$s.
In general, hybrid fitnesses will depend on the specific geometry of the optimal set.
As such, depending on the particular geometry, as well as the parental distances,
Haldane's rule might not always occur.
The key point we would like to communicate is that when $F_2$ hybrid breakdown \emph{can} occur
(i.e. due to geometry and distance),
and system coefficients are distributed across the autosomes and sex chromosomes,
Haldane's rule will manifest.
We have added some clarifying text \revref.

}

\begin{point}{}
  \llname{r21}: This paragraph summarises the main body of work performed in this paper. Thus, consider moving it to the end of the Introduction (i.e. switching places with the last paragraph).
\end{point}

\reply{Thank you for this suggestion, but we would prefer to keep the current paragraph order. 
We agree that the second-to-last paragraph does provide a nice summary of the work, however,
after trying out the alternate ordering,
we decided that the current order
(which first discusses system theory, followed by quantitative genetics)
is more consistent with the ordering of the rest of the text.
}

### 2021-08-11 review-responses2.tex

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

We are writing to submit a revision of our manuscript, ``{System drift and speciation}'' (ID 21-0262), for your review.
We'd would like to thank you and the reviewers for your enthusiasm
and your suggestions -- they have helped us improve our manuscript's accuracy and clarity.

Both reviewers made comments about the section on Haldane's rule, which we have revised accordingly.
We also made an effort to provide more context for our manuscript by citing the additional literature, as requested by Reviewer 1. 
To address some of Reviewer 2's concerns, we added more discussion and detail pertaining to system theory and its biological interpretation.

Line numbers in the response to reviewers refer to the revised document;
in addition we are submitting a color-coded pdf showing differences to the previous submission
that has its own copy of the review responses at the end (and line number references are consistent within this file as well).

We hope that you enjoy the resulting manuscript as much as we do.

\vskip 5em

### 2021-08-12 kalman_walk_paper.tex
cover letter
edits
fixed fig5 script
maybe add an inline recombine mech?

@@

\begin{figure}[H]
  %\centering
 % \begin{tabular}{ccc}
 %   \includegraphics[width=0.5\textwidth, height=0.125\paperheight]{examples/osc_F1_hundreth_tau0} &
 %     \includegraphics[width=0.5\textwidth, height=0.125\paperheight]{examples/osc_F2_hundrethtau0} \\
 %     \includegraphics[width=0.5\textwidth, height=0.125\paperheight]{examples/osc_F1_tenth_tau0.pdf} &
 %     \includegraphics[width=0.5\textwidth, height=0.125\paperheight]{examples/osc_F2_tenthtau0.pdf} \\
 %     \includegraphics[width=0.5\textwidth, height=0.125\paperheight]{examples/osc_F1_half_tau0.pdf} &
 %     \includegraphics[width=0.5\textwidth, height=0.125\paperheight]{examples/osc_F2_halftau0.pdf}
 % \end{tabular}
  %\includegraphics[scale=0.5]{examples/f1f2_newplot}
  \includegraphics[scale=0.5]{examples/blue_hybrids}
  \caption{
    \textbf{(left)} Phenotypes of $F_1$ hybrids between a homozygous $A(0)$ parent and, 
    top-to-bottom, homozygous $A(\sfrac{1}{100})$, $A(\sfrac{1}{10})$, and $A(\sfrac{1}{2})$ parents,
    where $A(\epsilon)$ is defined in Figure \ref{fig:all_osc};
    parental coefficients differ by around 0.5\%, 5\%, and 25\% respectively.
    Parental phenotypes ($\sin t + \cos t$) are shown in solid black, and hybrid phenotypes in blue.
    \textbf{(right)} Phenotypes of all $3^4 = 81$ possible $F_2$ hybrids between the same set of parents,
    with parental phenotype again in black.
    %% PLR: There are at most 128 possible F2s, not 256; but many are equivalent; 
    %%   it only matters if the offspring is homo/het/homozygous, so there are 3^4 = 81 possible expression patterns.
    Blue lines correspond to different $F_2$ hybrids,
    many of which show complete breakdown.
  } \label{fig:hybs}
\end{figure}

>>>>

\jss{Could add something like this?}

In Example \ref{ex:hybrid_osc} and Figure \ref{fig:hybs}, however, meiosis is modeled such that it recombines system coefficients independently,
to form haploid gametes. E.g., a possible oscillator gamete could be formed by shuffling the diploid's haploid genomes -- with one genome's coefficients notated with Greek and the other with Latin, as follows,
$\left[ \begin{smallmatrix} \alpha & \beta \\ \gamma & \delta \end{smallmatrix} \right] \times \left[ \begin{smallmatrix} a & b \\ c & d \end{smallmatrix} \right] \to \left[ \begin{smallmatrix} \alpha & b \\ c & \delta \end{smallmatrix} \right]$.

@@

\begin{figure}[H]
  \includegraphics{examples/blue_hybrids2}
  \caption{
    \textbf{(left)} Phenotypes of $F_1$ hybrids between a homozygous $A(0)$ parent and, 
    top-to-bottom, homozygous $A(\sfrac{1}{100})$, $A(\sfrac{1}{10})$, and $A(\sfrac{1}{2})$ parents,
    where $A(\epsilon)$ is defined in Figure \ref{fig:all_osc};
    parental coefficients differ by around 0.5\%, 5\%, and 25\% respectively.
    Parental phenotypes ($\sin t + \cos t$) are shown in solid black, and hybrid phenotypes in blue.
    \textbf{(right)} Phenotypes of all $3^4 = 81$ possible $F_2$ hybrids between the same set of parents,
    with parental phenotype again in black.
    %% PLR: There are at most 128 possible F2s, not 256; but many are equivalent; 
    %%   it only matters if the offspring is homo/het/homozygous, so there are 3^4 = 81 possible expression patterns.
    $F_2$ hybrids, many of which show complete breakdown, are shown in shades of blue and grey. 
  } \label{fig:hybs}
\end{figure}
